---
title: "CodeBook"
author: "Saloni Bhatia"
date: "August 7, 2019"
output: html_document
---
---

title: "Codebook.rmd"

author: "Saloni Bhatia"

date: "8/16/2019"

output: html_document

---



```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
install.packages("kableExtra")
install.packages("knitr")
install.packages("readxl")

library(knitr)

library(kableExtra)


```



## DDS Analytics Talent Management Study Codebook



This codebook provides an overview of the data as it was received and any modifications that were required to support our analysis. It is a documentation of the variables in the data set, any normalization or modifications that were made, and should support replication of our analysis. 



### Data Files Received



DDSAnalytics provided unique(Consulting) with a Microsoft Excel (.xlsx) file titled CaseStudy2-data.xlsx. 



```{r}

df <- read.csv("C:/SMU/DoingDatascience/Casestudy/CaseStudy2-data.csv")

```



### Initial Data Structure



The data contained within the provided table has the following variable names and types: 

<br>



```{r}
kable(t(lapply(df,class)))


#kable_styling(bootstrap_options = c("striped", "hover"))

#kable(t(lapply(cs2Raw,class)))%>%

 # kable_styling()%>%

 # scroll_box(height = 100, width = 500)

```



### Test Models



A number of test models were created to help identify features provided in the data set that contribute to Attrition rates within the population. 



#### Model 1: Logistic Regression



A logistic regression model to predict whether an employee -- based on provided features -- would be in class (AttritionCode = 1) or out of Class. This model accepts all of the provided categorical and continuous variables and determines their impact on Attrition. The model is un-normalized and unweighted and provides only a first step analysis. Despite this, however, it is capable of achieving accuracy rates of 87% with Precision values of 90.7% and Recall of 94.9%. 



This model is described by the statement: 



```{r eval=FALSE, echo=TRUE}

AttritionCode ~ Age + JobRole + StockOptionLevel + WorkLifeBalance + 

              YearsSinceLastPromotion + EnvironmentSatisfaction + JobInvolvement + 

              OverTime + RelationshipSatisfaction + YearsAtCompany + YearsWithCurrManager + 

              BusinessTravel + DistanceFromHome + Gender + MaritalStatus + 

              NumCompaniesWorked + YearsInCurrentRole

```